### 2020-03-04 main_text/main_text.tex
GC comments and unit formatting
Finished GC S1 Text comments. Small typos/formatting corrections.

	\subsection*{Materials and Methods}
	\paragraph*{Bacterial Strains.} Strains used in this work
	derive from {\it E. coli} K12 MG1655 with the {\it lac} operon
	removed and were constructed similarly to those used in
	\cite{RazoMejia2018, Belliveau2019}. $\Delta{xapAB}$ and
	$\Delta{xapABR}$ knockouts were generated using the approach of
	Datsenko and Wanner \cite{Datsenko2000}. For the fluorescent
	reporter constructs the {\it xapAB} promoter region, with one or
	both XapR binding sites, were extracted from {\it E. coli} K12
	MG1655 by PCR. These were cloned into the YFP-expressing reporter
	plasmid pZS25,  containing kanamycin resistance, and integrated at
	the {\it galK} locus using $\lambda$ Red recombineering \cite{Sharan2009}.

	\paragraph*{Flow Cytometry Measurements.}
	Experimental measurements in Fig. 2B of the main text were obtained
	via flow cytometry and were performed in a similar manner as
	described in~\cite{RazoMejia2018}. Briefly, wild-type,
	$\Delta{xapABR}$, and $\Delta{xapAB}$ were grown to saturation in
	rich LB Miller (BD Medical) medium and were diluted 1:1000 into M9
	minimal medium supplemented with 0.5\% (w/v) glucose and the
	appropriate concentration of xanthosine. For measurements of the
	$\Delta{xapAB}$ strain, chloramphenicol was also added to the growth
	medium to ensure stability of the pZ3PN25-tetR plasmid along with 10
	ng /mL of anhydro-tetracycline to maximize expression of
	xapR-mCherry. Cultures were allowed to grow at 37$^\circ$ C for six
	to eight hours to an OD$_{600nm}$ of $\approx$ 0.3. At this point,
	the cultures were diluted 1:10 into M9 + 0.5\% glucose and were
	immediately analyzed via a MacsQuant VYB Flow Cytometery
	(MiltenyiBiotec). Measurement files were exported to CSV file
	formats and analyzed as previously described
	in~\cite{RazoMejia2018}.

>>>>

	\subsection*{Materials and Methods}
	\paragraph*{Bacterial Strains.} Strains used in this work
	derive from {\it E. coli} K12 MG1655 with the {\it lac} operon
	removed and were constructed similarly to those used in~\cite{RazoMejia2018, Belliveau2019}. $\Delta{xapAB}$ and
	$\Delta{xapABR}$ knockouts were generated using the approach of
	Datsenko and Wanner~\cite{Datsenko2000}. For the fluorescent
	reporter constructs the {\it xapAB} promoter region, with one or
	both XapR binding sites, were extracted from {\it E. coli} K12
	MG1655 by PCR. These were cloned into the YFP-expressing reporter
	plasmid pZS25, containing kanamycin resistance, and integrated at
	the {\it galK} locus using $\lambda$ Red recombineering~\cite{Sharan2009}.

	\paragraph*{Flow Cytometry Measurements.}
	Experimental measurements in Fig. 2B of the main text were obtained
	via flow cytometry and were performed in a similar manner as
	described in~\cite{RazoMejia2018}. Briefly, wild-type,
	$\Delta{xapABR}$, and $\Delta{xapAB}$ were grown to saturation in
	rich LB Miller (BD Medical) medium and were diluted 1:1000 into M9
	minimal medium supplemented with 0.5\% (w/v) glucose and the
	appropriate concentration of xanthosine. For measurements of the
	$\Delta{xapAB}$ strain, chloramphenicol was also added to the growth
	medium to ensure stability of the pZ3PN25-tetR plasmid along with 10
	ng /mL of anhydro-tetracycline to maximize expression of
	xapR-mCherry. Cultures were allowed to grow at $37\unit{^\circ C}$ for six
	to eight hours to an OD$_{600\unit{nm}}$ of $\approx 0.3$. At this point,
	the cultures were diluted 1:10 into M9 + 0.5\% glucose and were
	immediately analyzed via a MacsQuant VYB Flow Cytometery
	(MiltenyiBiotec). Measurement files were exported to CSV file
	formats and analyzed as previously described
	in~\cite{RazoMejia2018}.

### 2020-03-04 s1_text/s1_text.tex
GC comments and unit formatting
Finished GC S1 Text comments. Small typos/formatting corrections.

\usepackage{placeins}
\usepackage{refcount}
\usepackage{textgreek}
\usepackage{bm}
\usepackage{url} % url in Quellen verlinken

@@

\begin{flushleft}
	{\Large
		\textbf\newline{Supporting information for ``theoretical investigation of a genetic switch for metabolic adaptation''}
	}
	\newline
	% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
		\\
	Kathrin S. Laxhuber\textsuperscript{1},
	Muir J. Morrison\textsuperscript{2},
	Griffin Chure\textsuperscript{3},
	Rob Phillips\textsuperscript{2,3$\ast$}
	\\
	\bigskip
	\textbf{1} Department of Chemistry and Applied Biosciences, ETH Zurich, 8093 Zurich, Switzerland
	\\
	\textbf{2} Department of Physics, California Institute of Technology, Pasadena, CA, USA
	\\
	\textbf{3} Department of Biology and Biological Engineering, California Institute of Technology, Pasadena, CA, USA
	\\
	\bigskip

@@

\paragraph*{Exponential protein decay.}
The protein decay should be dominated by dilution through cell division, which is a discrete process. However, the cell grows before division, resulting in a continuous dilution. Furthermore, replication leads to a varying number of gene copies and thereby also a varying transcription rate. Trying to capture all these mechanistic details would lead to an overcomplicated model. Treating the protein decay as a Poisson process has produced correct results in past research, which is why we expect this assumption to be reasonable.

\paragraph*{Kinetics of XapB, NupC, and NupG.}
Entering the cell consists of two steps. In a first transport step, xanthosine passes the
outer membrane through porins like Tsx, OmpF and OmpC. This does not seem to
be a rate-limiting step. Then, XapB actively transports it across the inner
membrane, powered by the proton motive force. There are two other active
nucleoside transporters that seem to be transporting xanthosine with a very
low affinity: NupC and NupG. It was found that \textDelta\emph{nupC}
\textDelta\emph{nupG} strains cannot grow on xanthosine. Hence, these seem
to be necessary to ``start'' the system by importing small amounts of
xanthosine that then activate \emph{xapB}, and there appears to be no other
significant way in which xanthosine can enter the cell. All statements in this paragraph are based on data
from~\cite{Norholm2001}.

@@

\section{Experimental materials and methods} \label{sec:exp}
\paragraph{Construction of $\Delta{xapABR}$ deletion strain.} Cells containing
plasmid pSIM6 were diluted 1:100 in 50 ml LB media and grown to an OD600
of $\approx 0.4$ at $30 ^\circ$C. The culture was immediately placed in
a water bath shaker at 43$^\circ$C for 15 minutes and then cooled in an
ice bath for 10 minutes. Cells were then spun down for 10 minutes
(4,000~$g$, 4$^\circ$C) and resuspended on ice in 50~ml of chilled
water. This was repeated three times before resuspending in 200~$\mu$L
of chilled water to generate competent cells. Homologous primer
extension sequences were obtained from Baba {\it et al.} \cite{Baba2006}
and used to generate linear DNA containing a kanamycin resistance
gene insert by PCR (amplified from plasmid pKD4), using primers that.
contained homology for the region on the chromosome to be deleted
\cite{Datsenko2000}. Electroporation was performed using 1~$\mu$L
purified PCR product (about 100~ng DNA), mixed with 50~$\mu$L cells.
Cells were resuspended in 750~$\mu$L SOC media and outgrowth was
performed on a shaker at 30$^\circ$C for 90-120 minutes. Cells were then
selected on an LB-agar plate containing kanamycin (30~$\mu$g/ml) and
grown overnight at 30$^\circ$C. The deletions were confirmed by both
colony PCR and sequencing. After confirmation, the deletion was
transferred to a clean strain through P1 transduction and selection on
kanamycin. Removal of kanamycin resistance was achieved by transforming
cells with plasmid pCP20 (expressing FLP recombinase, carbenicillin
resistance, on a temperature-sensitive replicon), with cells grown at
30$^\circ$C. Lastly, several colonies were then growth colony-purified
once non-selectively at 43$^\circ$C and then tested for loss of
antibiotic resistance for both kanamycin and carbenicillin.

\paragraph{Construction of $\Delta{xapAB}$ strain.}
Further experiments are underway in which precise control of the xapR copy number are needed and the preliminary data from these experiments are shown in the middle panel of Fig. 2A of the main text. This strain was constructed by creating a XapR-mCherry fusion protein expressed from a tetR regulated promoter. This fusion construct was integrated into the ycbN locus of the $\Delta{xapABR}$ strain making an effective $\Delta{xapAB}$ strain. A pZS3-PN25-TetR expression plasmid was also inserted into this genetic background which permits titratable control over the xapR-mCherry copy number. For the data presented in this work, expression was maximized using 10~ng/mL of the tetR inducer anhydro-tetracycline. Further information of a similar construct (LacI-mCherry) can be found in~\cite{Brewster2014} and~\cite{Chure2019}.

@@

>>>>

\usepackage{placeins}
\usepackage{refcount}
\usepackage{textgreek}
\usepackage{upgreek}
\usepackage{bm}
\usepackage{url} % url in Quellen verlinken

@@

\begin{flushleft}
	{\Large
		\textbf\newline{Supporting information for ``theoretical investigation of a genetic switch for metabolic adaptation''}
	}
	\newline
	% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
		\\
	Kathrin S. Laxhuber\textsuperscript{1},
	Muir J. Morrison\textsuperscript{2},
	Griffin Chure\textsuperscript{3},
	Rob Phillips\textsuperscript{2,3$\ast$}
	\\
	\bigskip
	\textbf{1} Department of Chemistry and Applied Biosciences, ETH Zurich, 8093 Zurich, Switzerland
	\\
	\textbf{2} Department of Physics, California Institute of Technology, Pasadena, CA, USA
	\\
	\textbf{3} Division of Biology and Biological Engineering, California Institute of Technology, Pasadena, CA, USA
	\\
	\bigskip

@@

\paragraph*{Exponential protein decay.}
The protein decay should be dominated by dilution through cell division, which is a discrete process. However, the cell grows before division, resulting in a continuous dilution in concentration. Furthermore, replication leads to a varying number of gene copies and thereby also a varying transcription rate. Trying to capture all these mechanistic details would lead to an overcomplicated model. Treating the protein decay as a Poisson process has produced correct results in past research, which is why we expect this assumption to be reasonable.

\paragraph*{Kinetics of XapB, NupC, and NupG.}
All statements in this paragraph are based on data
from~\cite{Norholm2001}. Xanthosine entering the cell consists of two steps. In a first transport step, xanthosine passes the
outer membrane through porins like Tsx, OmpF and OmpC. This does not seem to
be a rate-limiting step. Then, XapB actively transports it across the inner
membrane, powered by the proton motive force. There are two other active
nucleoside transporters that seem to be transporting xanthosine with a very
low affinity: NupC and NupG. It was found that \textDelta\emph{nupC}
\textDelta\emph{nupG} strains cannot grow on xanthosine. Hence, these seem
to be necessary to ``start'' the system by importing small amounts of
xanthosine that then activate \emph{xapB}, and there appears to be no other
significant way in which xanthosine can enter the cell.

@@

\section{Experimental materials and methods} \label{sec:exp}
\paragraph{Construction of $\bm{\Delta{xapABR}}$ deletion strain.} Cells containing
plasmid pSIM6 were diluted 1:100 in $50 \unit{mL}$ LB media and grown to an OD600
of $\approx 0.4$ at $30^\circ$C. The culture was immediately placed in
a water bath shaker at $43\unit{^\circ C}$ for 15 minutes and then cooled in an
ice bath for 10 minutes. Cells were then spun down for 10 minutes
($4,000 \unit{g}$, $4\unit{^\circ C}$) and resuspended on ice in $50\unit{mL}$ of chilled
water. This was repeated three times before resuspending in $200 \unit{\upmu L}$
of chilled water to generate competent cells. Homologous primer
extension sequences were obtained from Baba {\it et al.} \cite{Baba2006}
and used to generate linear DNA containing a kanamycin resistance
gene insert by PCR (amplified from plasmid pKD4), using primers that.
contained homology for the region on the chromosome to be deleted
\cite{Datsenko2000}. Electroporation was performed using $1 \unit{\upmu L}$
purified PCR product (about $100 \unit{ng}$ DNA), mixed with $50 \unit{\upmu L}$ cells.
Cells were resuspended in $750 \unit{\upmu L}$ SOC media and outgrowth was
performed on a shaker at $30\unit{^\circ C}$ for 90-120 minutes. Cells were then
selected on an LB-agar plate containing kanamycin ($30 \unit{\upmu g/mL}$) and
grown overnight at $30\unit{^\circ C}$. The deletions were confirmed by both
colony PCR and sequencing. After confirmation, the deletion was
transferred to a clean strain through P1 transduction and selection on
kanamycin. Removal of kanamycin resistance was achieved by transforming
cells with plasmid pCP20 (expressing FLP recombinase, carbenicillin
resistance, on a temperature-sensitive replicon), with cells grown at
$30\unit{^\circ C}$. Lastly, several colonies were then growth colony-purified
once non-selectively at $43\unit{^\circ C}$ and then tested for loss of
antibiotic resistance for both kanamycin and carbenicillin.

\paragraph{Construction of $\bm{\Delta{xapAB}}$ strain.}
Further experiments are underway in which precise control of the xapR copy number are needed and the preliminary data from these experiments are shown in the middle panel of Fig. 2A of the main text. This strain was constructed by creating a XapR-mCherry fusion protein expressed from a tetR regulated promoter. This fusion construct was integrated into the ycbN locus of the $\Delta{xapABR}$ strain making an effective $\Delta{xapAB}$ strain. A pZS3-PN25-TetR expression plasmid was also inserted into this genetic background which permits titratable control over the xapR-mCherry copy number. For the data presented in this work, expression was maximized using $10 \unit{ng/mL}$ of the tetR inducer anhydro-tetracycline. Further information of a similar construct (LacI-mCherry) can be found in~\cite{Brewster2014} and~\cite{Chure2019}.

@@

	\bibitem{Chure2019}
	Chure G, Kaczmarek ZA, Phillips R.
	\newblock Physiological Adaptability and Parametric Versatility in a Simple
	Genetic Circuit.
	\newblock bioRxiv. 2019;doi:{10.1101/2019.12.19.878462}.
